An Analysis of Health & Lifestyle Factors Associated with Sleep Quality

### 2024-04-19 products/manuscript/Manuscript.qmd
Linear Models Table
Added a table and description for the linear model results to the manuscript

Given that there were only 17 variables after final data transformation,
simple linear regression models were performed to determine baseline
associations between variables and Sleep Quality (the outcome of
interest). These regression models were created to determine which
variables had the largest impact on Sleep Quality and drove the rest of
the analyses. Simple linear regression models were fitted for the
following variables: BMI, cat_bp, Stress Level, Gender, Age, and
Occupation.

@@

We believed that stress and sleep quality would be negatively
correlated, sleep duration and quality would be positively correlated,
and that jobs with longer or more variable hours would negatively affect
sleep quality. We found that all of our assumptions were supported; the
results of our generalized linear models (GLMs) are included in the
supplementary material folder as there are too many of them to include
in the main manuscript.

@@

![The ranking of each variable's importance in determining sleep quality
according to our Random Forest model.](images/variableimportance.png)

@@

![Table 1: A comparison of the RMSE and R-squared values generated by
the null and RF models. The null model has no R-squared value due to its
method of generating predictions.](images/metric_table.png){width="510"}

As seen in table 1, our null model was found to have an RMSE of 1.193;
compared to our best parameter RF model's RMSE of 0.059, this is
relatively high. The R-squared value doesn't exist for the null model
since there is technically no variation between the observed and
predicted values.

@@

Through our linear models (included in the supplementary materials
folder), we determined that our hypotheses regarding each variable's
effect on quality of sleep were supported. The variables determined to
be the most important by our RF model had the following effects on sleep
quality: sleep duration was positively correlated, age was positively
correlated, and stress level was negatively correlated.

>>>>

Given that there were only 17 variables after final data transformation,
simple linear regression models were performed to determine baseline
associations between variables and Sleep Quality (the outcome of
interest). These regression models were created to determine which
variables had the largest impact on Sleep Quality and drove the rest of
the analyses. Simple linear regression models were fitted for the
following variables: BMI, cat_bp, Stress Level, Physical Activity Level,
Sleep Duration, Gender, Age, and Occupation. The other variables were
determined to have too weak of a relationship after our EDA and were
therefore left out of the analyses.

@@

![Table 1: The results of our GLMs. The "term" column identifies the
category/variable being analyzed. The "model" column identifies the
predictors used to fit each model. The "statistic" column is a relative
measure of the strength of the relationship between each predictor and
sleep quality. The "estimate" column shows the predicted change in sleep
quality score as a result of each associated
predictor.](images/table.png)

We believed that stress and sleep quality would be negatively
correlated, sleep duration and quality would be positively correlated,
and that jobs with longer or more variable hours would negatively affect
sleep quality. As seen in table 1, we found that all of our assumptions
were supported. The "intercept" values under the term column represents
what the expected sleep quality score would be if a continuous predictor
was at a value of 0. For categorical variables, it represents the
reference category (e.g. the intercept for the Gender model represents
the expected sleep quality score for females).

@@

![Figure 6: The ranking of each variable's importance in determining
sleep quality according to our Random Forest
model.](images/variableimportance.png)

@@

![Table 2: A comparison of the RMSE and R-squared values generated by
the null and RF models. The null model has no R-squared value due to its
method of generating predictions.](images/metric_table.png){width="510"}

As seen in table 2, our null model was found to have an RMSE of 1.193;
compared to our best parameter RF model's RMSE of 0.059, this is
relatively high. The R-squared value doesn't exist for the null model
since there is technically no variation between the observed and
predicted values.

@@

Through our linear models, we determined that our hypotheses regarding
each variable's effect on quality of sleep were supported. The variables
determined to be the most important by our RF model had the following
effects on sleep quality: sleep duration was positively correlated, age
was positively correlated, and stress level was negatively correlated.
